\documentclass[10pt]{armath}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{enumitem}

\title{Rendering Equation and BSDF}
\author{Arden Rasmussen}
\date{\today}

\newcommand{\lp}{\left(}
  \newcommand{\rp}{\right)}
\newcommand{\lv}{\left|}
\newcommand{\rv}{\right|}
\newcommand{\lvv}{\left\Vert}
\newcommand{\rvv}{\right\Vert}
\renewcommand{\S}{\mathcal{S}}
\renewcommand{\b}[1]{\mathbf{#1}}
\renewcommand{\vec}[1]{\overrightarrow{#1}}

\begin{document}
\maketitle

\section{Rendering Equation}%
\label{sec:rendering_equation}

\begin{align*}
  L_o\lp\b{p},
  \b{o}\rp=L_e\lp\b{p},\b{o}\rp+\int_{\S^2}f_s\lp\b{i},\b{o},\b{n}\rp
  L_i\lp\b{p},\b{i}\rp\lv\b{i}\cdot\b{n}\rv d\b{i}
\end{align*}

\begin{description}
  \item[$L_o$] The light going out of a point in a given direction.
  \item[$L_e$] The light emmited from a point in a given direction.
  \item[$Li$] The light comming into a point from a given direction.
  \item[$\b{p}$] Location of point in space.
  \item[$\S^2$] Is the unit sphere, containing all possible $\b{i}$.
  \item[$\Omega$] Is the unit hemisphere centered around $\b{n}$.
  \item[$\b{i}$] Is the direction of incident ray of light, for the most case
    this will be the negative of the direction along which we just ray
    traced/marched.
  \item[$\b{o}$] Is the negative direction of the outgoing ray oflight, this will be the direction along which we send new rays for tracing.
  \item[$\b{n}$] Is the surface normal at the point $\b{p}$.
  \item[$f_s$] Is the BSDF given from the objects material.
\end{description}

\section{BSDF}%
\label{sec:bsdf}

\begin{align*}
  f_s\lp\b{i},\b{o},\b{n}\rp=f_r\lp\b{i},\b{o},\b{n}\rp+f_t\lp\b{i},\b{o},\b{n}\rp
\end{align*}

\begin{align*}
  f_r\lp\b{i},\b{o},\b{n}\rp&=\frac{F\lp\b{i},\b{h_r}\rp
    G\lp\b{i},\b{o},\b{h_r}\rp
    D\lp\b{h_r}\rp}{4\lv\b{i}\cdot\b{n}\rv\lv\b{o}\cdot\b{n}\rv}\\
  f_t\lp\b{i},\b{o},\b{n}\rp&=\frac{\lv\b{i}\cdot\b{h_t}\rv\lv\b{o}\cdot\b{h_t}\rv}{\lv\b{i}\cdot\b{n}\rv\lv\b{o}\cdot\b{n}\rv}
  \frac{\eta_o^2\lp1-F\lp\b{i},\b{h_t}\rp\rp G\lp\b{i},\b{o},\b{h_t}\rp
    D\lp\b{h_t}\rp}{{\lp\eta_i\lp\b{i}\cdot\b{h_t}\rp+\eta_o\lp\b{o}\cdot\b{h_t}\rp\rp}^2}
\end{align*}
\begin{align*}
  \b{h_r}&=\frac{\b{i}+\b{o}}{\lvv\b{i}+\b{o}\rvv}\\
  \b{h_t}&=-\frac{\b{i}+\eta\b{o}}{\lvv\b{i}+\b{o}\rvv}
\end{align*}
\begin{description}
  \item[$\eta_o$] Is the index of refraction medium of the outgoing light.
  \item[$\eta_i$] Is the index of refraction medium of the incoming light.
  \item[$\eta$] is the ratio of the IORs of the two media,
    $\eta=\frac{\eta_o}{\eta_i}$. 
\end{description}

\end{document}
